---
title: "Untitled"
author: "Jordi Puigdellivol"
date: "8 de enero de 2016"
output: html_document
---

# Stadistical Test
```{r, echo=FALSE}
library(ggplot2)
```

## Introductio



In this document we will try to understand how, when different tests works and which one to use in different situations, also tries to make a document to explain it in a easy way through examples.
As long as this is a personal document, it can contain a lot of errors, be free to let mi know them ^^.



we will try to understant 3 diferentnt test, in a basic level and implement them in R.

 - [Kolmogorov-Smirnov test](https://en.wikipedia.org/wiki/Kolmogorov%E2%80%93Smirnov_test)
 - [Student's t-test](https://en.wikipedia.org/wiki/Student%27s_t-test)
 - [Chi-squared test](https://en.wikipedia.org/wiki/Chi-squared_test)

Also to exploit it's wekneses. We will test them in 5 diferent distributions or convinations of them to try to get wrong results.

 - [Normal distribution](https://en.wikipedia.org/wiki/Normal_distribution)
 - [Uniform distribution](https://en.wikipedia.org/wiki/Uniform_distribution_(continuous))
 - [Exponential distribution](https://en.wikipedia.org/wiki/Exponential_distribution)
 - [Log-normal distribution](https://en.wikipedia.org/wiki/Log-normal_distribution)
 - [Logistic distribution](https://en.wikipedia.org/wiki/Logistic_distribution)

all this distributions has parameters.

we will try to make a the test with in 3 diferent ways with the relation between parameters.

 - same paramaters
 - close parameters
 - very different parameters
 
so let's start!!!

# What is A/B testing

## A/B testing

[A/B testing](https://en.wikipedia.org/wiki/A/B_testing)

# Kolmogorov-Smirnov Tests

## ks.test()

```{r}

x <- rnorm(50)
y <- runif(30)
# Do x and y come from the same distribution?
ks.test(x, y)
# Does x come from a shifted gamma distribution with shape 3 and rate 2?
ks.test(x+2, "pgamma", 3, 2) # two-sided, exact
ks.test(x+2, "pgamma", 3, 2, exact = FALSE)
ks.test(x+2, "pgamma", 3, 2, alternative = "gr")

# test if x is stochastically larger than x2
x2 <- rnorm(50, -1)
plot(ecdf(x), xlim = range(c(x, x2)))
plot(ecdf(x2), add = TRUE, lty = "dashed")
t.test(x, x2, alternative = "g")
wilcox.test(x, x2, alternative = "g")
ks.test(x, x2, alternative = "l")
```

# Chi-squared Test

## chisq.test()

```{r}
## From Agresti(2007) p.39
M <- as.table(rbind(c(762, 327, 468), c(484, 239, 477)))
dimnames(M) <- list(gender = c("F", "M"),
                    party = c("Democrat","Independent", "Republican"))
(Xsq <- chisq.test(M))  # Prints test summary
Xsq$observed   # observed counts (same as M)
Xsq$expected   # expected counts under the null
Xsq$residuals  # Pearson residuals
Xsq$stdres     # standardized residuals


## Effect of simulating p-values
x <- matrix(c(12, 5, 7, 7), ncol = 2)
chisq.test(x)$p.value           # 0.4233
chisq.test(x, simulate.p.value = TRUE, B = 10000)$p.value
                                # around 0.29!

## Testing for population probabilities
## Case A. Tabulated data
x <- c(A = 20, B = 15, C = 25)
chisq.test(x)
chisq.test(as.table(x))             # the same
x <- c(89,37,30,28,2)
p <- c(40,20,20,15,5)
try(
chisq.test(x, p = p)                # gives an error
)
chisq.test(x, p = p, rescale.p = TRUE)
                                # works
p <- c(0.40,0.20,0.20,0.19,0.01)
                                # Expected count in category 5
                                # is 1.86 < 5 ==> chi square approx.
chisq.test(x, p = p)            #               maybe doubtful, but is ok!
chisq.test(x, p = p, simulate.p.value = TRUE)

## Case B. Raw data
x <- trunc(5 * runif(100))
chisq.test(table(x))            # NOT 'chisq.test(x)'!
```

# Student's t-Test

## t.test()
```{r}
require(graphics)

t.test(1:10, y = c(7:20))      # P = .00001855
t.test(1:10, y = c(7:20, 200)) # P = .1245    -- NOT significant anymore

## Classical example: Student's sleep data
plot(extra ~ group, data = sleep)
## Formula interface
t.test(extra ~ group, data = sleep)
ggplot(sleep)+geom_density(aes(x=extra,fill=group),alpha=0.3,position = "identity")

```


# Student's t-Test

## wilcox.test()
```{r}
wilcox.test()
```


```{r, echo=FALSE}
```

# Personal test groups
## Normal distribution

```{r normal_distribution}
n <- 100
norm_data <- data.frame(id=1:n,x=rnorm(n),x_=rnorm(n),y=rnorm(n,0.5,1.5),y_=rnorm(n,3,5))
norm_data_sort<- data.frame(id=1:n,x=cumsum( sort(norm_data$x) ),x_=cumsum(sort(norm_data$x_)),y=cumsum( sort(norm_data$y)),y_=cumsum( sort(norm_data$y_) ) )

ks.tests     <- sapply( norm_data[,-c(1)] , function(x){ ks.test( norm_data$x , x ) } )
t.tests      <- sapply( norm_data[,-c(1)] , function(x){ t.test ( norm_data$x , x ) } ) # con pocos datos, si no todo passa como normal
chisq.tests  <- sapply( norm_data[,-c(1)] , function(x){ chisq.test ( norm_data$x , x ,correct = F ) } )
wilcox.tests <- sapply( norm_data[,-c(1)] , function(x){ wilcox.test( norm_data$x , x  ) } )


ggplot( reshape2::melt(norm_data_sort,id.vars="id" ) ) + 
  geom_point(aes(x=id,y=value,colour=variable) )

ggplot(reshape2::melt(norm_data,id.vars="id" ))+
         geom_density(aes(x=value,fill=variable),alpha=0.3,position = "identity" )
```

```{r uniform_distribution}
n <- 100
unif_data <- data.frame(id=1:n,x=runif(n),x_=runif(n),y=runif(n,0.5,1.5),y_=runif(n,3,5))
unif_data_sort<- data.frame(id=1:n,x=cumsum( sort(unif_data$x) ),x_=cumsum(sort(unif_data$x_)),y=cumsum( sort(unif_data$y)),y_=cumsum( sort(unif_data$y_) ) )

ks.tests     <- sapply( unif_data[,-c(1:2)] , function(x){ ks.test( unif_data$x , x ) } )
t.tests      <- sapply( unif_data[,-c(1:2)] , function(x){ t.test ( unif_data$x , x ) } )
chisq.tests  <- sapply( unif_data[,-c(1:2)] , function(x){ chisq.test ( unif_data$x , x ,correct = F ) } )
wilcox.tests <- sapply( unif_data[,-c(1:2)] , function(x){ wilcox.test( unif_data$x , x  ) } )


ggplot( reshape2::melt(unif_data_sort,id.vars="id" ) ) + 
  geom_point(aes(x=id,y=value,colour=variable) )

ggplot(reshape2::melt(unif_data,id.vars="id" ))+
         geom_density(aes(x=value,fill=variable),alpha=0.3,position = "identity" )
```

```{r exponential_distribution}
n <- 100
exp_data <- data.frame(id=1:n,x=rexp(n),x_=rexp(n),y=rexp(n,1.5),y_=rexp(n,3))
exp_data_sort<- data.frame(id=1:n,x=cumsum( sort(exp_data$x) ),x_=cumsum(sort(exp_data$x_)),y=cumsum( sort(exp_data$y)),y_=cumsum( sort(exp_data$y_) ) )

ks.tests     <- sapply( exp_data[,-c(1:2)] , function(x){ ks.test( exp_data$x , x ) } )
t.tests      <- sapply( exp_data[,-c(1:2)] , function(x){ t.test ( exp_data$x , x ) } )
chisq.tests  <- sapply( exp_data[,-c(1:2)] , function(x){ chisq.test ( exp_data$x , x ,correct = F ) } )
wilcox.tests <- sapply( exp_data[,-c(1:2)] , function(x){ wilcox.test( exp_data$x , x  ) } )


ggplot( reshape2::melt(exp_data_sort,id.vars="id" ) ) + 
  geom_point(aes(x=id,y=value,colour=variable) )

ggplot(reshape2::melt(exp_data,id.vars="id" ))+
         geom_density(aes(x=value,fill=variable),alpha=0.3,position = "identity" )
```

```{r log-norm_distribution}
n <- 100
lnorm_data <- data.frame(id=1:n,x=rlnorm(n),x_=rlnorm(n),y=rlnorm(n,0.5,1.5),y_=rlnorm(n,3,2))
lnorm_data_sort<- data.frame(id=1:n,x=cumsum( sort(lnorm_data$x) ),x_=cumsum(sort(lnorm_data$x_)),y=cumsum( sort(lnorm_data$y)),y_=cumsum( sort(lnorm_data$y_) ) )

ks.tests     <- sapply( lnorm_data[,-c(1:2)] , function(x){ ks.test( lnorm_data$x , x ) } )
t.tests      <- sapply( lnorm_data[,-c(1:2)] , function(x){ t.test ( lnorm_data$x , x ) } )
chisq.tests  <- sapply( lnorm_data[,-c(1:2)] , function(x){ chisq.test ( lnorm_data$x , x ,correct = F ) } )
wilcox.tests <- sapply( lnorm_data[,-c(1:2)] , function(x){ wilcox.test( lnorm_data$x , x  ) } )


ggplot( reshape2::melt(lnorm_data_sort,id.vars="id" ) ) + 
  geom_point(aes(x=id,y=value,colour=variable) )

ggplot(reshape2::melt(lnorm_data,id.vars="id" ))+
         geom_density(aes(x=value,fill=variable),alpha=0.3,position = "identity")+coord_cartesian()

ggplot(reshape2::melt(lnorm_data,id.vars="id" ))+
         geom_density(aes(x=log(value),fill=variable),alpha=0.3,position = "identity")+coord_cartesian()
```

# References

http://www.physics.csbsju.edu/stats/KS-test.html
http://www.itl.nist.gov/div898/handbook/eda/section3/eda35g.htm
https://en.wikipedia.org/wiki/Kolmogorov%E2%80%93Smirnov_test
https://en.wikipedia.org/wiki/Student%27s_t-test
https://en.wikipedia.org/wiki/Chi-squared_test
